Deriving Multi-Headed Projective Dependency Parses \\ from Link Grammar Parses

### 2014-05-31 doc/LinksToDAG.tex
updated the sample english parse section of the appendix in the paper.

\begin{abstract}
Under multi-headed dependency grammar, a parse is a DAG rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
% OLD
% investigate the benefit of such parsers or to work on making them faster or more accurate.  
To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of \input{figure/conll_analysis_sentences_total.tex}parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.  We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.  
\end{abstract}

@@

For every encountered label we generate two boolean label/direction variables (e.g. label/left, label/right). Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints. The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite\cite{achterberg2009scip}.

@@

@@

\clearpage
\appendix
\section{ILP Formulation}
Here we will present our ILP formulation.

@@

\section{Sample English Parses}

\todo[inline]{Show a bunch of English sentences with CoNLL vs.\@ link parses.
  Explain how they were selected and how the results are colorized etc.}

Here we display a set of link parses as further examples

>>>>

\begin{abstract}
Under multi-headed dependency grammar, a parse is a directed acyclic graph rather than a tree.  Such formalisms can be more syntactically and semantically expressive.  However, 
it is hard to train, test, or improve multi-headed parsers because few multi-headed corpora exist, particularly for the projective case.
% OLD
% investigate the benefit of such parsers or to work on making them faster or more accurate.  
To help fill this gap, we observe that link grammar already produces {\em undirected} projective graphs.  
% OLD
% ... produces parses that are similar to multi-headed projective dependency parses, except that the links are undirected. 
We use Integer Linear Programming to assign consistent directions to the labeled links in a corpus of \input{figure/conll_analysis_sentences_total.tex}parses produced by the Link Grammar Parser, which has broad-coverage hand-written grammars of English as well as Russian and other languages.  We find that such directions can indeed be consistently assigned in a way that yields valid multi-headed dependency parses. The resulting parses in English appear reasonably linguistically plausible, although they are not in general consistent with CoNLL-style parses of the same sentences; we discuss the differences.  
\end{abstract}

@@

For every encountered label we generate two boolean label/direction variables (e.g. label/left, label/right). Setting one of these variables to \textsc{true} allows the label type to only go in the specified direction, while setting both would allow the label to go in either way.
The objective is to minimize the number of label/direction variables set to \textsc{true}, while still satisfying acyclicity and reachability constraints. The encoding was written in the Zimpl little language \cite{Koch2004} and solved using the SCIP Optimization Suite \cite{achterberg2009scip}.

@@

\section{English and Russian Data}
The data used for this work came from English and Russian corpuses. The English sentences were taken from the bnews conll training data set \noteJH{cite} and provided about 60,000 sentence examples in total.

The Russian sentence data was taken from a part of the Russian National Corpus \noteJH{cite?}

@@

\clearpage
\appendix
\section{ILP Formulation}
We present our ILP formulation.

@@

\section{Sample English Parses}
In this section, we will display 100 sample English parses. The top half of each parse visualization represents the CoNLL dependency parse, while the bottom half shows our directionalized link grammar parse.

\subsection{Sentence Selection}
These sentences have been taken from the English bnews training data set. We chose to display sentences with complete link parses similar to the analysis section of this paper. We display the first 100 parses whose character lengths do not exceed the width of the page.

\subsection{Parse Colorization}
The top blue arcs represent the CoNLL dependency parses. For every arc:
\begin{enumerate}
\item If there is a corresponding link edge between the same nodes then a solid blue line is drawn.
\item Or else a dotted blue line is drawn. 
\end{enumerate}
The bottom red arcs represent the directionalized link grammar parses. For every link edge:
\begin{enumerate}
\item If there is a corresponding CoNLL arc between the same nodes then a solid red line is drawn.
\item If the matching arc is facing the opposite direction, then the link edge is drawn thicker.
\item Of the remaining edges, if an edge is a multiheaded link, then it is drawn with thick orange.
\item The rest of the remaining edges are drawn with dotted red lines.
\end{enumerate}

\subsection{Link Parser Annotations}
The sentence used in the middle of the parse is taken from the output of the link parser. We removed sentences with dropped words, but they would have shown up in the parse as [bracketed]. There are also words that the link parser did not recognize, but rather made a guess as to its part of speech. These words are tagged with a [!] symbol at the end. The dashes in the link grammar part of speech tag annotations mean that the link parser did not return one.

% Uncomment this to add in all of the parses. Warning: compilation will take a long time.
%\input{figure/parses.tikz}
